### 2018-07-06 Overview.Rmd
Inital exploration of data

---
title: "Dataset Summary"
author: "Daniel Falster & Susie Zajitschek"
date: "06/07/2018"
output: 
  html_document:
    fig_height: 8
    fig_width: 10
    df_print: paged
    rows.print: 25
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(readr)
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
```

Here is an initial exploration of the mouse data we are working on.

The data is big, but not so large that we can't use our standard tools on a desktop. I'd suggest using packages from the [tidyverse](https://www.tidyverse.org/) family, in particular `readr`, `dplyr`, `ggplot2`, `skimr`, `scales`.

##  Dataset overview

Read in data, specifying variable types:

```{r cars}
data_raw <- read_csv("data/dr7.0_all_control_data.csv", 
                  col_types = cols(
                      .default = col_character(),
                      project_id = col_character(),
                      id = col_character(),
                      parameter_id = col_character(),
                      age_in_days = col_integer(),
                      date_of_experiment = col_datetime(format = ""),
                      weight = col_double(),
                      phenotyping_center_id = col_character(),
                      production_center_id = col_character(),
                      weight_date = col_datetime(format = ""),
                      date_of_birth = col_datetime(format = ""),
                      procedure_id = col_character(),
                      pipeline_id = col_character(),
                      biological_sample_id = col_character(),
                      biological_model_id = col_character(),
                      weight_days_old = col_integer(),
                      datasource_id = col_character(),
                      experiment_id = col_character(),
                      data_point = col_double(),
                      age_in_weeks = col_integer(),
                      `_version_` = col_character()
                      )
                )
```

Number of rows:

```{r}
data_raw %>% nrow()
```

Variables in the dataset:

```{r}
data_raw %>% names()
```

Lots of variables are full of NAs

```{r}
n_records <- data_raw %>% nrow()
NA_count <- data_raw %>%
  summarise_all(funs(sum(is.na(.))/n_records*100)) %>% 
  unlist() %>% sort(decreasing=TRUE)
(NA_count)

# make a plot
tibble(percent_NA = NA_count) %>% 
  ggplot(aes(percent_NA)) + geom_histogram()
```

Let's remove those variables that have all NAs and create a new variable called data:

```{r}
(all_NAs <- names(NA_count)[NA_count == 100])
data <- data_raw %>% select(-one_of(all_NAs))
```

Now we `r data %>% names() %>% length()` variables in the dataset:

```{r}
data %>% names()
```

Now use `skimr` to take a quick look of all variables:

```{r}
x <- data %>% skimr::skim()
x %>% skimr::kable()
```

Next we'll look at some specific variables of potential importance.

## Production center

Contributions by `production_center`:
```{r}
data$production_center %>% table() %>% sort()
```

## Weights

Overall distribution of weights:
```{r}
ggplot(data, aes(x=weight)) + 
  geom_histogram(bins=50)
```

Weights by center:
```{r}
ggplot(data, aes(x=weight)) + 
  geom_histogram() + 
  facet_wrap( ~ production_center, ncol=1)
```

Weights by center and sex:
```{r}
ggplot(data, aes(x=weight, fill=sex)) + 
  geom_histogram(bins=50) + 
  facet_wrap( ~ production_center, ncol=1)
```

## Procedure

Contributions by `procedure_name`:

```{r, fig_height=12}
data$procedure_name %>% table() %>% sort(decreasing = TRUE)

x <- data %>% group_by(procedure_name) %>% summarise(n=n())
ggplot(x, aes(reorder(procedure_name, n), n)) +
  geom_col() + coord_flip()
```

Procedure name by production center:

```{r}
table(data$production_center, data$procedure_name)
```

Next want to visualise the above, but so far the plot is unmanagable (need to improve):

```{r, fig.height=25}

x <- data %>% 
  group_by(production_center, procedure_name) %>% 
  summarise(n=n())

ggplot(x, aes(reorder(production_center, n), n)) +
  geom_col() + coord_flip() +
  facet_wrap( ~ procedure_name, ncol=4)
```