### 2018-08-03 Cleaning & subset.Rmd
Procedures for cleaning and exporting a subset

---
title: "Dataset Summary"
author: "Daniel Falster"
date: "06/07/2018"
output: 
  html_document:
    fig_height: 6
    fig_width: 10
    df_print: paged
    rows.print: 10
    code_folding: hide
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(readr)
library(dplyr)
```

# Load

```{r cars}
data_raw <- read_csv("data/dr7.0_all_control_data.csv", 
                  col_types = cols(
                      .default = col_character(),
                      project_id = col_character(),
                      id = col_character(),
                      parameter_id = col_character(),
                      age_in_days = col_integer(),
                      date_of_experiment = col_datetime(format = ""),
                      weight = col_double(),
                      phenotyping_center_id = col_character(),
                      production_center_id = col_character(),
                      weight_date = col_datetime(format = ""),
                      date_of_birth = col_datetime(format = ""),
                      procedure_id = col_character(),
                      pipeline_id = col_character(),
                      biological_sample_id = col_character(),
                      biological_model_id = col_character(),
                      weight_days_old = col_integer(),
                      datasource_id = col_character(),
                      experiment_id = col_character(),
                      data_point = col_double(),
                      age_in_weeks = col_integer(),
                      `_version_` = col_character()
                      )
                )
```

# Cleaning:

```{r}
data <- data_raw %>% 
  mutate(parameter_name = tolower(parameter_name) ) %>%
  filter(age_in_days > 0 & age_in_days < 500) %>% 
  select(production_center, project_name, biological_sample_id, sex, procedure_name, age_in_days, weight, parameter_name, data_point) %>% 
  arrange(production_center, biological_sample_id, age_in_days)
```

# Export

Body size only

```{r}
dir.create("export", F, F)
data %>% filter(parameter_name == "body weight") %>%
  saveRDS("export/body_weight.rds")
```