### 2018-08-03 Cleaning & subset.Rmd
Add strain into body size data; summary stats by population

---
title: "Dataset Summary"
author: "Daniel Falster"
date: "06/07/2018"
output: 
  html_document:
    fig_height: 6
    fig_width: 10
    df_print: paged
    rows.print: 10
    code_folding: hide
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(readr)
library(dplyr)
```

# Load

```{r cars}
data_raw <- read_csv("data/dr7.0_all_control_data.csv", 
                  col_types = cols(
                      .default = col_character(),
                      project_id = col_character(),
                      id = col_character(),
                      parameter_id = col_character(),
                      age_in_days = col_integer(),
                      date_of_experiment = col_datetime(format = ""),
                      weight = col_double(),
                      phenotyping_center_id = col_character(),
                      production_center_id = col_character(),
                      weight_date = col_datetime(format = ""),
                      date_of_birth = col_datetime(format = ""),
                      procedure_id = col_character(),
                      pipeline_id = col_character(),
                      biological_sample_id = col_character(),
                      biological_model_id = col_character(),
                      weight_days_old = col_integer(),
                      datasource_id = col_character(),
                      experiment_id = col_character(),
                      data_point = col_double(),
                      age_in_weeks = col_integer(),
                      `_version_` = col_character()
                      )
                )
```

# Cleaning:

```{r}
data <- data_raw %>% 
  mutate(parameter_name = tolower(parameter_name) ) %>%
  filter(age_in_days > 0 & age_in_days < 500) %>% 
  select(production_center, strain_name, experiment_id, external_sample_id, biological_sample_id, sex, procedure_name, age_in_days, weight, parameter_name, data_point) %>% 
  arrange(production_center, biological_sample_id, age_in_days)
```

# Export

Body size only

```{r}
dir.create("export", F, F)
data %>% filter(parameter_name == "body weight") %>%
  saveRDS("export/body_weight.rds")
```

# For variance analysis

- population has > 5 individuals

keep age closet t0 10 
```{r}

data %>% 
  filter(
    age_in_days >= 98, 
    parameter_name == "body weight") %>%
  mutate(age_diff=abs(100-age_in_days)) %>%
  group_by(biological_sample_id) %>%
  filter(age_diff == min(age_diff)) %>%
  mutate(population = sprintf("%s-%s-%s", strain_name, production_center, sex) )  -> x

i <- match(unique(x$biological_sample_id), x$biological_sample_id)

data2 <- x[i,] 
```

Calculate number of reocrds per population
```{r}
data2 %>% 
  group_by(population) %>%
  summarise(
    n_records = n_distinct(biological_sample_id)
  ) -> population_size
hist(population_size$n_records)

sum(population_size$n_records < 10)
```

```{r}

data %>% 
  filter(
    age_in_days >= 100, 
    parameter_name == "body weight") %>%
  select(biological_sample_id, external_sample_id) %>%
  arrange(external_sample_id) %>%
  data.frame()

data %>% 
  filter(
    age_in_days >= 100, 
    parameter_name == "body weight") %>%
  mutate(age_diff=abs(100-age_in_days)) %>%
  group_by(biological_sample_id) %>%
  filter(age_diff == min(age_diff)) -> x

i <- match(unique(x$biological_sample_id), x$biological_sample_id)

x[i,] -> x

# check only one record per individual

table(x$biological_sample_id) -> z 
z[z!=1]

```