### 2018-08-03 Cleaning & subset.Rmd
Meta analysis of body size at ~100 days

---
title: "Dataset Summary"
author: "Daniel Falster"
date: "06/07/2018"
output: 
  html_document:
    fig_height: 6
    fig_width: 10
    df_print: paged
    rows.print: 10
    code_folding: hide
    theme: yeti
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
library(readr)
library(dplyr)
```

# Load

```{r cars}
data_raw <- read_csv("data/dr7.0_all_control_data.csv", 
                  col_types = cols(
                      .default = col_character(),
                      project_id = col_character(),
                      id = col_character(),
                      parameter_id = col_character(),
                      age_in_days = col_integer(),
                      date_of_experiment = col_datetime(format = ""),
                      weight = col_double(),
                      phenotyping_center_id = col_character(),
                      production_center_id = col_character(),
                      weight_date = col_datetime(format = ""),
                      date_of_birth = col_datetime(format = ""),
                      procedure_id = col_character(),
                      pipeline_id = col_character(),
                      biological_sample_id = col_character(),
                      biological_model_id = col_character(),
                      weight_days_old = col_integer(),
                      datasource_id = col_character(),
                      experiment_id = col_character(),
                      data_point = col_double(),
                      age_in_weeks = col_integer(),
                      `_version_` = col_character()
                      )
                )
```

# Cleaning:

```{r}
data <- data_raw %>% 
  mutate(parameter_name = tolower(parameter_name) ) %>%
  filter(age_in_days > 0 & age_in_days < 500) %>% 
  select(production_center, strain_name, experiment_id, external_sample_id, 
         biological_sample_id, sex, procedure_name, age_in_days, weight, parameter_name, data_point) %>% 
  arrange(production_center, biological_sample_id, age_in_days)
```

Export for body size only

```{r}
dir.create("export", F, F)
data %>% filter(parameter_name == "body weight") %>%
  saveRDS("export/body_weight.rds")
```

# For variance analysis

- population has > 5 individuals

keep age closet t=100

```{r}

data %>% 
  filter(
    age_in_days >= 98, 
    parameter_name == "body weight") %>%
  mutate(age_diff=abs(100-age_in_days)) %>%
  group_by(biological_sample_id) %>%
  filter(age_diff == min(age_diff)) %>%
  mutate(population = sprintf("%s-%s-%s", strain_name, production_center, sex) )  -> x

# still some individuals with multiple records (bc appear under different procedures, so filter to one record)
i <- match(unique(x$biological_sample_id), x$biological_sample_id)
data2 <- x[i,] 
```

Calculate number of reocrds per population

```{r}
data2 %>% 
  group_by(population) %>%
  summarise(
    n_records = n_distinct(biological_sample_id)
  ) -> population_size
hist(population_size$n_records)

sum(population_size$n_records < 5)
```

Now calculate statsics by population:

```{r}
population_stats <- data2 %>% 
  filter(!is.na(data_point)) %>% 
  group_by(population, sex, strain_name, production_center) %>% 
  summarise(
    trait = parameter_name[1],
    x_bar = mean(data_point),
    x_sd = sd(data_point),
    n_ind = n()
  ) %>% 
  ungroup() %>%
  filter(n_ind > 5) %>% 
  # Check both sexes present & filter those missing
  group_by(strain_name, production_center) %>% 
  mutate(
    n_sex = n_distinct(sex)
  ) %>% ungroup() %>%
  filter(n_sex ==2) %>% 
  arrange(strain_name, production_center, sex)
```

Now analyse by population data:

Sanity checks

```{r}
i <- seq(1, nrow(population_stats), by=2)
all(population_stats$sex[i] == "female")
all(population_stats$sex[i+1] == "male")
all(population_stats$strain_name[i+1] == population_stats$strain_name[i] )
all(population_stats$production_center[i+1] == population_stats$production_center[i] )

input <- data.frame(
       n1i = population_stats$n_ind[i], 
       n2i = population_stats$n_ind[i+1], 
       x1i = population_stats$x_bar[i], 
       x2i = population_stats$x_bar[i+1],
       sd1i = population_stats$x_sd[i], 
       sd2i = population_stats$x_sd[i+1])

all(input > 0)
```
Check plot of mean vs variance

```{r}
plot(population_stats$x_bar, population_stats$x_sd, pch=16, log="xy", col = as.factor(population_stats$sex))
```

```{r}
Calc.lnCVR <- function(CMean, CSD, CN, EMean, ESD, EN){
	S2 <- log(ESD) - log(EMean) + 1 / (2*(EN - 1)) - (log(CSD) - log(CMean) + 1 / (2*(CN - 1)))
  return(S2)
}

Calc.var.lnCVR <- function(CMean, CSD, CN, EMean, ESD, EN, Equal.E.C.Corr=T) {
	if(Equal.E.C.Corr==T){
		mvcorr <- cor.test(log(c(CMean, EMean)), log(c(CSD, ESD)))$estimate
		S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * mvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * mvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))
	}
	else{
		Cmvcorr <- cor.test(log(CMean), log(CSD))$estimate
		Emvcorr <- cor.test(log(EMean), (ESD))$estimate
		S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * Cmvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * Emvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))		
	}
	return(S2)
}

Calc.lnVR <- function(CSD, CN, ESD, EN){
	log(ESD) - log(CSD) + 1 / (2*(EN - 1)) - 1 / (2*(CN - 1))
}

Calc.var.lnVR <- function( CN,  EN) {
	1 / (2*(EN - 1)) + 1 / (2*(CN - 1))
}

Calc.lnRR <- function(CMean, CSD, CN, EMean, ESD, EN) {
	log(EMean) - log(CMean)
}

Calc.var.lnRR <- function(CMean, CSD, CN, EMean, ESD, EN) {
	CSD^2/(CN * CMean^2) + ESD^2/(EN * EMean^2)
}
```

```{r}
# library(metafor)
#measures: CVR, VR
# escalc(measure="CVR", 
#       n1i = n1i, n2i = n2i, x1i = x1i, x2i = x2i, sd1i = sd1i, sd2i = sd2i, data = input)
# not working

results <- population_stats[i,] %>% 
  select(strain_name, production_center) %>%
  mutate(
    effect_size_CVR = Calc.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
    sample_variance_CVR = Calc.var.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
    effect_size_VR = Calc.lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
    sample_variance_VR = Calc.var.lnVR(CN = input$n1i, EN = input$n2i),
    effect_size_RR = Calc.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
    sample_variance_RR = Calc.var.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
    err = as.factor(seq_len(n()))
    )

```

## Compare CVR

```{r}
plot(results$effect_size_CVR, 1/sqrt(results$sample_variance_CVR), ylab="sample precision", xlab="effect size", pch=16, xlim=c(-0.7, 0.7), col=as.factor(results$production_center))
abline(v=0)
```

```{r}
metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR, random = list(~1| strain_name, ~1|production_center, ~1|err), data = results)
```

## Compare VR

```{r}
plot(results$effect_size_VR, 1/sqrt(results$sample_variance_VR), ylab="sample precision", xlab="effect size", pch=16, xlim=c(-0.8, 0.8), col=as.factor(results$production_center))
abline(v=0)
```

```{r}
metafor::rma.mv(yi = effect_size_VR, V = sample_variance_VR, random = list(~1| strain_name, ~1|production_center, ~1|err), data = results)
```

## Compare Means

```{r}
plot(results$effect_size_RR, 1/sqrt(results$sample_variance_RR), ylab="sample precision", xlab="effect size", pch=16, xlim=c(-0.5, 0.5), col=as.factor(results$production_center))
abline(v=0)
```

```{r}
metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR, random = list(~1| strain_name, ~1|production_center, ~1|err), data = results)
```