### 2019-07-25 R/Heterogeneity.Rmd
Add files via upload. Need to reconcile with Heterogeneity script, which will probably now become part of the main ms, not supplement

---
title: "IMPC Mouse data - Heterogeneity"
author: "Susanne & Felix Zajitschek"
date: "4rth April 2019"
output:
  html_document:
   code_folding: hide
   toc: true
   toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

```

```{r, include=FALSE}
getwd()
```

# Packages

```{r}
library(readr)
library(dplyr)
library(metafor)
library(devtools)
library(purrr)
library(tidyverse)
library(tibble)
library(kableExtra)
library(robumeta)
library(ggpubr)
library(ggplot2)
```
####################################################################################################################################

# ** FIGURE 4 (Second-order meta analysis on heterogeneity)

# Loading functions that are necessary
##Preparation of raw data
1) Data loading and cleaning of the csv file
This step we have already done and provide a the cleaned up file which is less computing intensive. However, cvs
```{r}

# loads the raw data, setting some default types for various columns

load_raw <- function(filename) {
  read_csv(filename, 
                              col_types = cols(
                                .default = col_character(),
                                project_id = col_character(),
                                id = col_character(),
                                parameter_id = col_character(),
                                age_in_days = col_integer(),
                                date_of_experiment = col_datetime(format = ""),
                                weight = col_double(),
                                phenotyping_center_id = col_character(),
                                production_center_id = col_character(),
                                weight_date = col_datetime(format = ""),
                                date_of_birth = col_datetime(format = ""),
                                procedure_id = col_character(),
                                pipeline_id = col_character(),
                                biological_sample_id = col_character(),
                                biological_model_id = col_character(),
                                weight_days_old = col_integer(),
                                datasource_id = col_character(),
                                experiment_id = col_character(),
                                data_point = col_double(),
                                age_in_weeks = col_integer(),
                                `_version_` = col_character()
                              )
  )
}

# Apply some standard cleaning to the data
clean_raw_data <- function(mydata) {
  mydata %>%

    # Fileter to IMPC source (recommened by Jeremey in email to Susi on 20 Aug 2018)
    filter(datasource_name == 'IMPC') %>%

    # standardise trait names
    mutate(parameter_name = tolower(parameter_name) ) %>%

    # remove extreme ages
    filter(age_in_days > 0 & age_in_days < 500) %>%

    # remove NAs 
    filter(!is.na(data_point)) %>%

    # subset to reasonable set of variables
    # date_of_experiment: Jeremy suggested using as an indicator of batch-level effects
    select(production_center, strain_name, strain_accession_id, biological_sample_id, pipeline_stable_id, procedure_group, procedure_name, sex, date_of_experiment, age_in_days, weight, parameter_name, data_point) %>% 
    arrange(production_center, biological_sample_id, age_in_days)
}

```

2) Subsetting the data to choose only one datapoint per individual per trait
```{r}
# this is a necessary step for the loop across all traits 
data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min, age_center) {
  tmp <- mydata %>%
    filter(age_in_days >= age_min,
           id == parameter) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)

  # still some individuals with multiple records (because same individual appears under different procedures, so filter to one record)
  j <- match(unique(tmp$biological_sample_id), tmp$biological_sample_id)
  tmp[j, ] 
}
```

## Functions for preparing the data for meta analyses
3) "Population statistics"
```{r}
calculate_population_stats <- function(mydata, min_individuals = 5) {
  mydata %>% 
    group_by(population, strain_name, production_center, sex) %>% 
    summarise(
      trait = parameter_name[1],
      x_bar = mean(data_point),
      x_sd = sd(data_point),
      n_ind = n()
    ) %>% 
    ungroup() %>%
    filter(n_ind > min_individuals) %>% 
    # Check both sexes present & filter those missing
    group_by(population) %>% 
    mutate(
      n_sex = n_distinct(sex)
    ) %>% 
    ungroup() %>%
    filter(n_sex ==2) %>% 
    select(-n_sex) %>%
    arrange(production_center, strain_name, population, sex)
}
```

 4) Extraction of effect sizes and sample variances
```{r}
create_meta_analysis_effect_sizes <- function(mydata) {
  i <- seq(1, nrow(mydata), by = 2)
  input <- data.frame(
    n1i = mydata$n_ind[i],
    n2i = mydata$n_ind[i + 1],
    x1i = mydata$x_bar[i],
    x2i = mydata$x_bar[i + 1],
    sd1i = mydata$x_sd[i],
    sd2i = mydata$x_sd[i + 1]
  )

  mydata[i,] %>% 
  select(strain_name, production_center) %>%
    mutate(
      effect_size_CVR = Calc.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_CVR = Calc.var.lnCVR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      effect_size_VR = Calc.lnVR(CSD = input$sd1i, CN = input$n1i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_VR = Calc.var.lnVR(CN = input$n1i, EN = input$n2i),
      effect_size_RR = Calc.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      sample_variance_RR = Calc.var.lnRR(CMean = input$x1i, CSD = input$sd1i, CN = input$n1i, EMean = input$x2i, ESD = input$sd2i, EN = input$n2i),
      err = as.factor(seq_len(n()))
    )

}
```

 5) Meta Analysis

 Function to calculate metanalaysis statistics. Created by A M Senior @ the University of Otago NZ 03/01/2014

Below are funcitons for calculating effect sizes for meta-analysis of variance. 
All functions take the mean, sd and n from the control and experimental groups.

The first function, Cal.lnCVR, calculates the the log repsonse-ratio of the coefficient of variance (lnCVR) - see Nakagawa et al 2015.

The second function calculates the measuremnt error variance for lnCVR. As well as the aforementioned parameters, this function also takes
Equal.E.C.Corr (default = T), which must be True or False. If true, the funciton assumes that the correlaiton between mean and sd (Taylor's Law) 
is equal for the mean and control groups, and, thus these data are pooled. If False the mean-SD correlation for the experimental and control groups
are calculated seperatley from one another.

Similar functions are then implemented for lnVR (for comparison of standard deviations) and ln RR  (for comparison of means)

```{r}

Calc.lnCVR <- function(CMean, CSD, CN, EMean, ESD, EN){
  log(ESD) - log(EMean) + 1 / (2*(EN - 1)) - (log(CSD) - log(CMean) + 1 / (2*(CN - 1)))
}

Calc.var.lnCVR <- function(CMean, CSD, CN, EMean, ESD, EN, Equal.E.C.Corr=T) {
  if(Equal.E.C.Corr==T){
    mvcorr <- cor.test(log(c(CMean, EMean)), log(c(CSD, ESD)))$estimate
    S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * mvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * mvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))
  }
  else{
    Cmvcorr <- cor.test(log(CMean), log(CSD))$estimate
    Emvcorr <- cor.test(log(EMean), (ESD))$estimate
    S2 <- CSD^2 / (CN * (CMean^2)) + 1 / (2 * (CN - 1)) - 2 * Cmvcorr * sqrt((CSD^2 / (CN * (CMean^2))) * (1 / (2 * (CN - 1)))) + ESD^2 / (EN * (EMean^2)) + 1 / (2 * (EN - 1)) - 2 * Emvcorr * sqrt((ESD^2 / (EN * (EMean^2))) * (1 / (2 * (EN - 1))))		
  }
  S2
}

Calc.lnVR <- function(CSD, CN, ESD, EN){
  log(ESD) - log(CSD) + 1 / (2*(EN - 1)) - 1 / (2*(CN - 1))
}

Calc.var.lnVR <- function( CN,  EN) {
  1 / (2*(EN - 1)) + 1 / (2*(CN - 1))
}

Calc.lnRR <- function(CMean, CSD, CN, EMean, ESD, EN) {
  log(EMean) - log(CMean)
}

Calc.var.lnRR <- function(CMean, CSD, CN, EMean, ESD, EN) {
  CSD^2/(CN * CMean^2) + ESD^2/(EN * EMean^2)
}

```

Having loaded the necessary functions, we can get started on the dataset.

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can ve loaded and cleaned using the da cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than th e .csv - which we can still provide for those who absoliutely want to start from scratch?)
```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")

data1 <- readRDS("export/data_clean.rds")

```

## Clean data
This requires the selection of traits that have been measured in at least 2 centers. Consequently, rare or unusual methods and procedures are being filtered out in this step.
```{r }
dat1 <-
  data1 %>%
  group_by(parameter_name) %>%
  summarize(center_per_trait = length(unique(production_center, na.rm = TRUE)))

dat2 <- merge(data1, dat1) 
dat_moreThan1center <-
  dat2  %>%
  filter(center_per_trait >= 2)

data2 <- dat_moreThan1center
#min(data2$center_per_trait)  # as a check ifthere indeed are no single occurences

```

#Preparation for Meta analysis
## Define population variable & add grouping variables

In this step, a grouping variable  is added (found in "Parameter.Grouping.csv")
The grouping variables were decided based on functional groups and procedures

```{r }
data3 <- data2 %>%
mutate(population = sprintf("%s-%s", production_center, strain_name))

 group <- read.csv("ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)]

```

## Assign each unique parameter_name (=trait,use trait variable) a unique number ('id')
We add a new variable, where redundant traits are combined
[note however, at this stage the dataset still contains non-sensical traits, i.e. traits that may not contain any u=information on variance]

```{r}
head(data)

names(data)[16] <- "parameter_group"

data <- transform(data, id = match(parameter_name, unique(parameter_name)))
n1 <- length(unique(data$parameter_name)) #232
n2 <- length(unique(data$parameter_group)) #161
n3 <- length(unique(data$procedure_name)) # 26

n <- length(unique(data$id))
#n  # just to check that the number of traits is 232
```

## *Create matrix to store results for all traits
```{r}
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*30), ncol = 30))) 
names(results.allhetero.grouping) <- c("id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR", 
	"s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR", 
	"s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR", 
	"s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", 
	"lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se")

```

# **LOOP
#  parameters to extract from metafor (sigma2's, s.nlevels)

```{r}

for(t in 1:n) {

  tryCatch({

    data_par_age <- data_subset_parameterid_individual_by_age(data, t, age_min = 0, age_center = 100)

    population_stats <- calculate_population_stats(data_par_age)

    results <- create_meta_analysis_effect_sizes(population_stats)

    # lnCVR, logaritm of the ratio of male and female coefficients of variance

    cvr. <- metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR, random = list(~1| strain_name, ~1|production_center,    
                                ~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), data = results)
    results.allhetero.grouping[t, 2] <- cvr.$sigma2[1]
    results.allhetero.grouping[t, 3] <- cvr.$sigma2[2]
    results.allhetero.grouping[t, 4] <- cvr.$sigma2[3]
    results.allhetero.grouping[t, 5] <- cvr.$s.nlevels[1]
    results.allhetero.grouping[t, 6] <- cvr.$s.nlevels[2]
    results.allhetero.grouping[t, 7] <- cvr.$s.nlevels[3]
     results.allhetero.grouping[t, 20] <- cvr.$b
     results.allhetero.grouping[t, 21] <- cvr.$ci.lb
     results.allhetero.grouping[t, 22] <- cvr.$ci.ub
     results.allhetero.grouping[t, 23] <- cvr.$se

    # lnVR, male to female variability ratio (logarithm of male and female standard deviations)

    vr. <- metafor::rma.mv(yi = effect_size_VR, V = sample_variance_VR, random = list(~1| strain_name, ~1|production_center,    
                                ~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), data = results)
    results.allhetero.grouping[t, 8] <- vr.$sigma2[1]
    results.allhetero.grouping[t, 9] <- vr.$sigma2[2]
    results.allhetero.grouping[t, 10] <- vr.$sigma2[3]
    results.allhetero.grouping[t, 11] <- vr.$s.nlevels[1]
    results.allhetero.grouping[t, 12] <- vr.$s.nlevels[2]
    results.allhetero.grouping[t, 13] <- vr.$s.nlevels[3]
     results.allhetero.grouping[t, 24] <- vr.$b
     results.allhetero.grouping[t, 25] <- vr.$ci.lb
     results.allhetero.grouping[t, 26] <- vr.$ci.ub
     results.allhetero.grouping[t, 27] <- vr.$se

    # lnRR, response ratio (logarithm of male and female means)

    rr. <- metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR, random = list(~1| strain_name, ~1|production_center,    
                                ~1|err), control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), data = results)
    results.allhetero.grouping[t, 14] <- rr.$sigma2[1]
    results.allhetero.grouping[t, 15] <- rr.$sigma2[2]
    results.allhetero.grouping[t, 16] <- rr.$sigma2[3]
    results.allhetero.grouping[t, 17] <- rr.$s.nlevels[1]
    results.allhetero.grouping[t, 18] <- rr.$s.nlevels[2]
    results.allhetero.grouping[t, 19] <- rr.$s.nlevels[3]
     results.allhetero.grouping[t, 28] <- rr.$b
     results.allhetero.grouping[t, 29] <- rr.$ci.lb
     results.allhetero.grouping[t, 30] <- rr.$ci.ub
     results.allhetero.grouping[t, 31] <- rr.$se

  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

```

## *Exclude traits
```{r}
results.allhetero.grouping2 <- results.allhetero.grouping[results.allhetero.grouping$s.nlevels.strain.VR != 0, ]
nrow(results.allhetero.grouping2) #218
```

# *Merge data sets containing metafor results with procedure etc. names 
```{r}
procedures <- read.csv("procedures.csv")

results.allhetero.grouping2$parameter_group <- data$parameter_group[match(results.allhetero.grouping2$id, data$id)]
results.allhetero.grouping2$procedure <- data$procedure_name[match(results.allhetero.grouping2$id, data$id)]

results.allhetero.grouping2$GroupingTerm <-  procedures$GroupingTerm[match(results.allhetero.grouping2$procedure, procedures$procedure)]
results.allhetero.grouping2$parameter_name <- data$parameter_name[match(results.allhetero.grouping2$id, data$id)]

```

## *Correlated parameters

```{r}
metahetero1 <- results.allhetero.grouping2 
length(unique(metahetero1$procedure)) #18
length(unique(metahetero1$GroupingTerm)) #9 
length(unique(metahetero1$parameter_group)) # 149 levels: one more tha in effect size alanlysis, see above; CHECK!
length(unique(metahetero1$parameter_name)) #218

```

# **Count of number of parameter names (correlated sub-traits) in each parameter group (par_group_size)

```{r}
metahetero1b <-
  metahetero1 %>%
  group_by(parameter_group) %>% 
  mutate(par_group_size = n_distinct(parameter_name))

metahetero1$par_group_size <- metahetero1b$par_group_size[match(metahetero1$parameter_group, metahetero1b$parameter_group)]

```

# **Create subsets with > 1 count (par_group_size > 1)

```{r}
metahetero1_sub <- subset(metahetero1, par_group_size > 1) # 90 observations  
str(metahetero1_sub)
# metahetero1_sub$sampleSize <- as.numeric(metahetero1_sub$sampleSize) #from previous analysis? don't think is used: : delete in final version
```

# **Nest data
```{r}
n_count. <- metahetero1_sub %>% 
  group_by(parameter_group) %>% 
  #mutate(raw_N = sum(sampleSize)) %>%  #don't think is necessary: delete in final version
  nest()

```

```{r}
model_count. <- n_count. %>% 
  mutate(model_lnRR = map(data, ~ robu(.x$lnRR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnRR_se)^2 )),
  model_lnVR = map(data, ~ robu(.x$lnVR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnVR_se)^2 )),
  model_lnCVR = map(data, ~ robu(.x$lnCVR ~ 1, data= .x, studynum= .x$id, modelweights = c("CORR"), rho = 0.8, 
                                                small = TRUE, var.eff.size= (.x$lnCVR_se)^2 )))

```
#Robumeta object details:
#str(model_count.$model_lnCVR[[1]])

## *Perform meta-analyses on correlated sub-traits, using robumeta

# Shinichi: We think we want to use these for further analyses:
# residual variance: as.numeric(robu_fit$mod_info$term1)     (same as 'mod_info$tau.sq')
# sample size: robu_fit$N

## **Extract and save parameter estimates 
```{r}
count_fun. <- function(mod_sub)
  return(c(as.numeric(mod_sub$mod_info$term1), mod_sub$N) )

```

```{r}
robusub_RR. <- model_count. %>% 
  transmute(parameter_group, estimatelnRR = map(model_lnRR, count_fun.)) %>% 
  mutate(r = map(estimatelnRR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnRR) %>%
  purrr::set_names(c("parameter_group","var.RR","N.RR"))

robusub_CVR. <- model_count. %>% 
  transmute(parameter_group, estimatelnCVR = map(model_lnCVR, count_fun.)) %>% 
  mutate(r = map(estimatelnCVR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnCVR) %>%
  purrr::set_names(c("parameter_group","var.CVR","N.CVR"))

robusub_VR. <- model_count. %>% 
  transmute(parameter_group, estimatelnVR = map(model_lnVR, count_fun.)) %>% 
  mutate(r = map(estimatelnVR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnVR) %>%
  purrr::set_names(c("parameter_group","var.VR","N.VR"))

robu_all. <- full_join(robusub_CVR., robusub_VR.) %>% full_join(., robusub_RR.)

```

## **Merge the two data sets (the new [robu_all] and the initial [uncorrelated sub-traits with count = 1]) 
##  In this step, we 	
1) put the N from robumeta and the  N from metafor (s.nlevels.error) together into the same columns (N.RR, N.VR, N.CVR)

2) calculate the total variance for metafor models as the sum of random effect variances and the residual error,
							then put them in the same columns together with the residual variances from robumeta
```{r}
metahetero_all <- metahetero1 %>% filter(par_group_size == 1) %>% as_tibble
metahetero_all$N.RR <- metahetero_all$s.nlevels.error.RR 
metahetero_all$N.CVR <- metahetero_all$s.nlevels.error.CVR 
metahetero_all$N.VR <- metahetero_all$s.nlevels.error.VR

# log (srt( of the data to look at))
metahetero_all$var.RR <- log(sqrt(metahetero_all$sigma2_strain.RR + metahetero_all$sigma2_center.RR + metahetero_all$sigma2_error.RR))
metahetero_all$var.VR <- log(sqrt(metahetero_all$sigma2_strain.VR + metahetero_all$sigma2_center.VR + metahetero_all$sigma2_error.VR))
metahetero_all$var.CVR <- log(sqrt(metahetero_all$sigma2_strain.CVR + metahetero_all$sigma2_center.CVR + metahetero_all$sigma2_error.CVR))
#str(metahetero_all)
#str(robu_all.)

#adjust the -inf to smallest number in dataset
metahetero_all <- metahetero_all %>% mutate(var.RR = if_else(var.RR == -Inf, -7 ,var.RR),
                   	  var.VR = if_else(var.VR == -Inf, -5 ,var.VR),
				  var.CVR = if_else(var.CVR == -Inf, -6 ,var.CVR))

```

# **Combine data 
# Step1 
```{r}
combinedmetahetero <- bind_rows(robu_all., metahetero_all)
#glimpse(combinedmetahetero)
```
# Steps 2&3
```{r}
metacombohetero <- combinedmetahetero
metacombohetero$counts <- metahetero1$par_group_size[match( metacombohetero$parameter_group, metahetero1$parameter_group)]
metacombohetero$procedure2 <-metahetero1$procedure[match( metacombohetero$parameter_group, metahetero1$parameter_group)]
metacombohetero$GroupingTerm2 <-metahetero1$GroupingTerm[match( metacombohetero$parameter_group, metahetero1$parameter_group)]

```
# **Clean-up and rename 
```{r}
metacombohetero <- metacombohetero[, c(1:7, 43:45)] 
names(metacombohetero)[9] <- "procedure" 
names(metacombohetero)[10] <- "GroupingTerm"

```

```{r}
#write.csv(metacombohetero, "metacombohetero.csv")

```

## *Last step: meta-meta-analysis of heterogeneity
```{r}
glimpse(metacombohetero)

## *Perform meta-meta-analysis (3 for each of the 9 grouping terms: var.CVR, var.VR, var.RR)

metacombohetero_final <- metacombohetero %>%
 group_by(GroupingTerm) %>%
 nest()

# **Calculate number of parameters per grouping term : not needed; delete in final version
#metacombo_final <- metacombo_final  %>%  mutate(para_per_GroupingTerm = map_dbl(data, nrow))

# **Final fixed effects meta-analyses within grouping terms, with SE of the estimate 
# Shinichi: meta-analyses on heterogeneity, using  sei as below, in metafor formula

heterog1 <- metacombohetero_final %>%

  mutate(model_heteroCVR = map(data, ~ metafor::rma.uni(yi = .x$var.CVR, sei = sqrt(1 / 2*(.x$N.CVR - 1)),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 10000, stepadj=0.5), verbose=F)),
	   model_heteroVR = map(data, ~ metafor::rma.uni(yi = .x$var.VR, sei = sqrt(1 / 2*(.x$N.VR - 1)),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 10000, stepadj=0.5), verbose=F)),
	   model_heteroRR = map(data, ~ metafor::rma.uni(yi = .x$var.RR, sei = sqrt(1 / 2*(.x$N.RR - 1)),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 10000, stepadj=0.5), verbose=F)))

heterog1

# **Re-structure data for each grouping term; extract heterogenenity/variance terms; delete un-used variables

Behaviour. <- heterog1 %>% filter(., GroupingTerm == "Behaviour")  %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

Immunology. <- heterog1 %>% filter(., GroupingTerm == "Immunology") %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

Hematology. <- heterog1 %>% filter(., GroupingTerm == "Hematology") %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

Hearing. <- heterog1 %>% filter(., GroupingTerm == "Hearing")  %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

Physiology. <- heterog1 %>% filter(., GroupingTerm == "Physiology")  %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

Metabolism. <- heterog1 %>% filter(., GroupingTerm == "Metabolism")  %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

Morphology. <- heterog1 %>% filter(., GroupingTerm == "Morphology") %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

Heart. <- heterog1 %>% filter(., GroupingTerm == "Heart")  %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

Eye. <- heterog1 %>% filter(., GroupingTerm == "Eye")  %>% select(., -data)  %>%  mutate(heteroCVR=.[[2]][[1]]$b, heteroCVR_lower=.[[2]][[1]]$ci.lb, heteroCVR_upper=.[[2]][[1]]$ci.ub, heteroCVR_se=.[[2]][[1]]$se,
	              heteroVR=.[[3]][[1]]$b, heteroVR_lower=.[[3]][[1]]$ci.lb, heteroVR_upper=.[[3]][[1]]$ci.ub, heteroVR_se=.[[3]][[1]]$se,
		        heteroRR=.[[4]][[1]]$b, heteroRR_lower=.[[4]][[1]]$ci.lb, heteroRR_upper=.[[4]][[1]]$ci.ub, heteroRR_se=.[[4]][[1]]$se) %>%
			  select(., GroupingTerm, heteroCVR:heteroRR_se)

heterog2 <- bind_rows(Behaviour., Morphology., Metabolism., Physiology., Immunology., Hematology., Heart., Hearing., Eye.)
#str(heterog2)

```

####################################################################################################################################
# *PLOTS

# **Restructure data for plotting 
```{r}
heterog3 <- gather(heterog2, parameter, value, c(heteroCVR, heteroVR, heteroRR), factor_key= TRUE)

heteroCVR.ci <- heterog3 %>% filter(parameter == "heteroCVR") %>% mutate(ci.low = heteroCVR_lower, ci.high = heteroCVR_upper)
heteroVR.ci <- heterog3 %>% filter(parameter == "heteroVR") %>% mutate(ci.low = heteroVR_lower, ci.high = heteroVR_upper)
heteroRR.ci <- heterog3 %>% filter(parameter == "heteroRR") %>% mutate(ci.low = heteroRR_lower, ci.high = heteroRR_upper)

heterog4 <- bind_rows(heteroCVR.ci, heteroVR.ci, heteroRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

# **Re-order grouping terms

heterog4$GroupingTerm <- factor(heterog4$GroupingTerm, levels =c("Behaviour","Morphology","Metabolism","Physiology","Immunology","Hematology","Heart","Hearing","Eye") )
heterog4$GroupingTerm <- factor(heterog4$GroupingTerm, rev(levels(heterog4$GroupingTerm)))
heterog4$label <- "All traits"
#write.csv(heterog4, "heterog4.csv")

```

# *Plot FIGURE 4 (Second-order meta analysis on heterogeneity)

# **Plot Fig4 all traits
```{r}

Metameta_Fig4_alltraits <- heterog4 %>%

  ggplot(aes(y= GroupingTerm, x= value)) +
  geom_errorbarh(aes(xmin = ci.low, 
                     xmax = ci.high), 
			   height = 0.1, show.legend = FALSE) +
  geom_point(aes(shape = parameter,
             fill = parameter, color = parameter), size = 2.2, 
             show.legend = FALSE) +
  scale_x_continuous(limits=c(-7.0, 1), 
                     #breaks = c(-2.0, -1.5, -1.0, -0.5, 0, 0.5, 1.0, 1.5, 2.0), 
                     name='Effect size') +
  facet_grid(cols = vars(parameter), rows = vars(label),
             labeller = label_wrap_gen(width = 23),
             scales= 'free', 
             space='free')+
  theme_bw()+
  theme(strip.text.y = element_text(angle = 270, size = 10, margin = margin(t=15, r=15, b=15, l=15)), 
	  strip.text.x = element_text(size = 12),
        strip.background = element_rect(colour = NULL, linetype = "blank", fill = "gray90"),
        text = element_text(size = 14),
        panel.spacing = unit(0.5, "lines"),
        panel.border= element_blank(),
        axis.line=element_line(), 
        panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(), 
        legend.title = element_blank(),
        axis.title.x = element_blank(),
	  axis.title.y = element_blank())

Metameta_Fig4_alltraits

```

######## Not sure whether the following makes sense for Fig 4 heterogeneity!!! ###########################

# *Prepare data for traits with CI not overlapping 0
```{r}
#create column with 1= different from zero, 0= zero included in CI

meta.plot4.sig <- metacombohetero %>% 
    	mutate(sigCVR = ifelse(heteroCVR_lower * heteroCVR_upper > 0, 1, 0),
	sigVR = ifelse(heteroVR_lower * heteroVR_upper > 0, 1, 0),
	sigRR = ifelse(heteroRR_lower * heteroRR_upper > 0, 1, 0))

#Significant subset for heteroCVR
metacombo_plot4.CVR <- meta.plot4.sig %>%
 filter(sigCVR == 1) %>%
 group_by(GroupingTerm) %>%
 nest()

#Significant subset for heteroVR
metacombo_plot4.VR <- meta.plot4.sig %>%
 filter(sigVR == 1) %>%
 group_by(GroupingTerm) %>%
 nest()

#Significant subset for heteroRR
metacombo_plot4.RR <- meta.plot4.sig %>%
 filter(sigRR == 1) %>%
 group_by(GroupingTerm) %>%
 nest()

# **Final fixed effects meta-analyses within grouping terms, with SE of the estimate

plot4.meta.CVR <- metacombo_plot4.CVR %>% 
  mutate(model_heteroCVR = map(data, ~ metafor::rma.uni(yi = .x$heteroCVR, sei = (.x$heteroCVR_upper - .x$heteroCVR_lower)/ 2*1.96,  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

plot4.meta.VR <- metacombo_plot4.VR %>% 
  mutate(model_heteroVR = map(data, ~ metafor::rma.uni(yi = .x$heteroVR, sei = (.x$heteroVR_upper - .x$heteroVR_lower)/ 2*1.96,  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

plot4.meta.RR <- metacombo_plot4.RR %>% 
  mutate(model_heteroRR = map(data, ~ metafor::rma.uni(yi = .x$heteroRR, sei = (.x$heteroRR_upper - .x$heteroRR_lower)/ 2*1.96,  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

# **Re-structure data for each grouping term; delete un-used variables

plot4.meta.CVR.b <- as.data.frame(plot4.meta.CVR %>% group_by(GroupingTerm) %>%  
		mutate(heteroCVR = map_dbl(model_heteroCVR, pluck(2)), heteroCVR_lower = map_dbl(model_heteroCVR, pluck(6)), 
		heteroCVR_upper =map_dbl(model_heteroCVR, pluck(7)), heteroCVR_se =map_dbl(model_heteroCVR, pluck(3))) )[, c(1,4:7)] 
add.row.hearing <- as.data.frame(t(c("Hearing", NA, NA, NA, NA))) %>% setNames(names(plot4.meta.CVR.b))
plot4.meta.CVR.b <- bind_rows(plot4.meta.CVR.b, add.row.hearing) 
plot4.meta.CVR.b <- plot4.meta.CVR.b[order(plot4.meta.CVR.b$GroupingTerm),]

plot4.meta.VR.b <- as.data.frame(plot4.meta.VR %>% group_by(GroupingTerm) %>%  
	mutate(heteroVR = map_dbl(model_heteroVR, pluck(2)), heteroVR_lower = map_dbl(model_heteroVR, pluck(6)), 
	heteroVR_upper =map_dbl(model_heteroVR, pluck(7)), heteroVR_se =map_dbl(model_heteroVR, pluck(3))) )[, c(1,4:7)] 
plot4.meta.VR.b <- plot4.meta.VR.b[order(plot4.meta.VR.b$GroupingTerm),]

plot3.meta.RR.b <- as.data.frame(plot3.meta.RR %>% group_by(GroupingTerm) %>%  
	mutate(lnRR = map_dbl(model_lnRR, pluck(2)), lnRR_lower = map_dbl(model_lnRR, pluck(6)), 
	lnRR_upper =map_dbl(model_lnRR, pluck(7)), lnRR_se =map_dbl(model_lnRR, pluck(3))) )[, c(1,4:7)] 
plot3.meta.RR.b <- plot3.meta.RR.b[order(plot3.meta.RR.b$GroupingTerm),]

overall.plot4 <- inner_join(plot4.meta.CVR.b, plot4.meta.VR.b) 
overall.plot4 <- inner_join(overall.plot4, plot4.meta.RR.b)

overall.plot4$GroupingTerm <- factor(overall.plot4$GroupingTerm, levels =c("Behaviour","Morphology","Metabolism","Physiology","Immunology","Hematology","Heart","Hearing","Eye") )
overall.plot4$GroupingTerm <- factor(overall.plot4$GroupingTerm, rev(levels(overall.plot4$GroupingTerm)))

```

# **Restructure data for plotting 
```{r}
overall4.sig <- gather(overall.plot4, parameter, value, c(heteroCVR, heteroVR, heteroRR), factor_key= TRUE)

heteroCVR.ci <- overall4.sig  %>% filter(parameter == "heteroCVR") %>% mutate(ci.low = heteroCVR_lower, ci.high = heteroCVR_upper)
heteroVR.ci <- overall4.sig  %>% filter(parameter == "heteroVR") %>% mutate(ci.low = heteroVR_lower, ci.high = heteroVR_upper)
heteroRR.ci <- overall4.sig  %>% filter(parameter == "heteroRR") %>% mutate(ci.low = heteroRR_lower, ci.high = heteroRR_upper)

overall4.sig <- bind_rows(heteroCVR.ci, heteroVR.ci, heteroRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

overall4.sig$label <- "CI not overlapping zero"

```

# **Plot Fig4 all significant results (CI not overlapping zero)
```{r}
######

Metameta_Fig4_sig <- overall4.sig %>%
  ggplot(aes(y= GroupingTerm, x= value)) +
  geom_errorbarh(aes(xmin = ci.low, 
                     xmax = ci.high), 
			   height = 0.1, show.legend = FALSE) +
  geom_point(aes(shape = parameter,
             fill = parameter, color = parameter), size = 2.2, 
             show.legend = FALSE) +
  scale_x_continuous(limits=c(-0.52, 0.4), 
                     breaks = c(-0.4, -0.2, 0, 0.2, 0.4), 
                     name='Effect size') +
  geom_vline(xintercept=0, 
             color='black', 
             linetype='dashed')+
  facet_grid(cols = vars(parameter), rows = vars(label),
             labeller = label_wrap_gen(width = 23),
             scales= 'free', 
             space='free')+
  theme_bw()+
  theme(strip.text.y = element_text(angle = 270, size = 10, margin = margin(t=15, r=15, b=15, l=15)), 
	  strip.text.x = element_blank(),
        strip.background = element_rect(colour = NULL, linetype = "blank", fill = "gray90"),
        text = element_text(size = 14),
        panel.spacing = unit(0.5, "lines"),
        panel.border= element_blank(),
        axis.line=element_line(), 
        panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(), 
        legend.title = element_blank(),
        axis.title.x = element_blank(),
	  axis.title.y = element_blank())

Metameta_Fig4_sig

#######################

# *Fig3 >10%

###########

########

# *Prepare data for traits with m/f difference > 10%
```{r}
#create column with 1= larger, 0= diff not larger than 10%

meta.plot3.perc <- metacombo %>% 
    	mutate(percCVR = ifelse(lnCVR < log (9/10) | lnCVR > log (11/10), 1, 0),
	percVR = ifelse(lnVR < log (9/10) | lnVR > log (11/10), 1, 0),
	percRR = ifelse(lnRR < log (9/10) | lnRR > log (11/10), 1, 0))

#Significant subset for lnCVR
metacombo_plot3.CVR.perc <- meta.plot3.perc %>%
 #filter(percCVR == 1) %>%
 group_by(GroupingTerm) %>%
 nest()

#Significant subset for lnVR
metacombo_plot3.VR.perc <- meta.plot3.perc %>%
 filter(percVR == 1) %>%
 group_by(GroupingTerm) %>%
 nest()

#Significant subset for lnRR
metacombo_plot3.RR.perc <- meta.plot3.perc %>%
 filter(percRR == 1) %>%
 group_by(GroupingTerm) %>%
 nest()

# **Final fixed effects meta-analyses within grouping terms, with SE of the estimate

plot3.meta.CVR.perc <- metacombo_plot3.CVR.perc %>% 
  mutate(model_lnCVR = map(data, ~ metafor::rma.uni(yi = .x$lnCVR, sei = (.x$lnCVR_upper - .x$lnCVR_lower)/ 2*1.96,  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

plot3.meta.VR.perc <- metacombo_plot3.VR.perc %>% 
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ 2*1.96,  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

plot3.meta.RR.perc <- metacombo_plot3.RR.perc %>% 
  mutate(model_lnRR = map(data, ~ metafor::rma.uni(yi = .x$lnRR, sei = (.x$lnRR_upper - .x$lnRR_lower)/ 2*1.96,  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

# **Re-structure data for each grouping term; delete un-used variables: "Hearing missing for all 3 parameters"

plot3.meta.CVR.perc.b <- as.data.frame(plot3.meta.CVR.perc %>% group_by(GroupingTerm) %>%  
		mutate(lnCVR = map_dbl(model_lnCVR, pluck(2)), lnCVR_lower = map_dbl(model_lnCVR, pluck(6)), 
		lnCVR_upper =map_dbl(model_lnCVR, pluck(7)), lnCVR_se =map_dbl(model_lnCVR, pluck(3))) )[, c(1,4:7)] 
add.row.hearing <- as.data.frame(t(c("Hearing", NA, NA, NA, NA))) %>% setNames(names(plot3.meta.CVR.perc.b))
plot3.meta.CVR.perc.b <- rbind(plot3.meta.CVR.perc.b, add.row.hearing) 
plot3.meta.CVR.perc.b <- plot3.meta.CVR.perc.b[order(plot3.meta.CVR.perc.b$GroupingTerm),]

plot3.meta.VR.perc.b <- as.data.frame(plot3.meta.VR.perc %>% group_by(GroupingTerm) %>%  
	mutate(lnVR = map_dbl(model_lnVR, pluck(2)), lnVR_lower = map_dbl(model_lnVR, pluck(6)), 
	lnVR_upper =map_dbl(model_lnVR, pluck(7)), lnVR_se =map_dbl(model_lnVR, pluck(3))) )[, c(1,4:7)] 
add.row.hearing <- as.data.frame(t(c("Hearing", NA, NA, NA, NA))) %>% setNames(names(plot3.meta.VR.perc.b))
plot3.meta.VR.perc.b <- rbind(plot3.meta.VR.perc.b, add.row.hearing) 
plot3.meta.VR.perc.b <- plot3.meta.VR.perc.b[order(plot3.meta.VR.perc.b$GroupingTerm),]

plot3.meta.RR.perc.b <- as.data.frame(plot3.meta.RR.perc %>% group_by(GroupingTerm) %>%  
	mutate(lnRR = map_dbl(model_lnRR, pluck(2)), lnRR_lower = map_dbl(model_lnRR, pluck(6)), 
	lnRR_upper =map_dbl(model_lnRR, pluck(7)), lnRR_se =map_dbl(model_lnRR, pluck(3))) )[, c(1,4:7)] 
add.row.hearing <- as.data.frame(t(c("Hearing", NA, NA, NA, NA))) %>% setNames(names(plot3.meta.RR.perc.b))
plot3.meta.RR.perc.b <- rbind(plot3.meta.RR.perc.b, add.row.hearing) 
plot3.meta.RR.perc.b <- plot3.meta.RR.perc.b[order(plot3.meta.RR.perc.b$GroupingTerm),]

overall.plot3.perc <- inner_join(plot3.meta.CVR.perc.b, plot3.meta.VR.perc.b) 
overall.plot3.perc <- inner_join(overall.plot3.perc, plot3.meta.RR.perc.b)

overall.plot3.perc$GroupingTerm <- factor(overall.plot3.perc$GroupingTerm, levels =c("Behaviour","Morphology","Metabolism","Physiology","Immunology","Hematology","Heart","Hearing","Eye") )
overall.plot3.perc$GroupingTerm <- factor(overall.plot3.perc$GroupingTerm, rev(levels(overall.plot3.perc$GroupingTerm)))

```

# **Restructure data for plotting 
```{r}
overall3.perc <- gather(overall.plot3.perc, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.perc <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

overall4.perc$label <- "Sex difference in m/f ratios > 10%"

overall4.perc$value <- as.numeric(overall4.perc$value)
overall4.perc$ci.low  <- as.numeric(overall4.perc$ci.low)
overall4.perc$ci.high <- as.numeric(overall4.perc$ci.high)

```

# **Plot Fig3 all >10% difference 
```{r}
######

Metameta_Fig3_perc <- overall4.perc %>%
  ggplot(aes(y= GroupingTerm, x= value)) +
  geom_errorbarh(aes(xmin = ci.low, 
                     xmax = ci.high), 
			   height = 0.1, show.legend = FALSE) +
  geom_point(aes(shape = parameter,
             fill = parameter, color = parameter), size = 2.2, 
             show.legend = FALSE) +
  scale_x_continuous(limits=c(-0.52, 0.4), 
                     breaks = c(-0.4, -0.2, 0, 0.2, 0.4), 
                     name='Effect size') +
  geom_vline(xintercept=0, 
             color='black', 
             linetype='dashed')+
  facet_grid(cols = vars(parameter), rows = vars(label),
             labeller = label_wrap_gen(width = 23),
             scales= 'free', 
             space='free')+
  theme_bw()+
  theme(strip.text.y = element_text(angle = 270, size = 10, margin = margin(t=15, r=15, b=15, l=15)), 
	  strip.text.x = element_blank(),
        strip.background = element_rect(colour = NULL, linetype = "blank", fill = "gray90"),
        text = element_text(size = 14),
        panel.spacing = unit(0.5, "lines"),
        panel.border= element_blank(),
        axis.line=element_line(), 
        panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(), 
        legend.title = element_blank(),
        axis.title.x = element_text(hjust = 0.5, size = 14),
	  axis.title.y = element_blank())

Metameta_Fig3_perc
```
#######################
# *Plot Fig3 all plots together
```{r}

library(ggpubr)
Fig3 <- ggarrange(Metameta_Fig3_alltraits, Metameta_Fig3_sig, Metameta_Fig3_perc, nrow = 3, align = "v", heights = c(1.22,1,1))

ggsave("Fig3.pdf", plot = Fig3, width = 6, height = 5)

####################################################################################################################################